### 2019-01-18 writing/thesis/chapters/results.tex
Changes.

\chapter{Results}\label{chap:results}
We discussed various scenarios in the previous chapter.
In this chapter, we use them to gauge the quality of our numerical scheme.
We start with a convergence test, followed by two standard testing scenarios.
Next, we simulate two different cloud scenarios and check, whether our \amr{} criterion is useful.

@@

\section{Accurate CFD}
After we established that our scheme converges, we evaluate it using standard \textsc{cfd} testing scenarios that either have an analytical solution or for which we have a reference solution.
% Show:
% \begin{itemize}
% \item Taylor-Green High Order
%   LIC, Plot v/p ana vs.\ real
% \item ABC Medium Order
%   Plot p ana.\ vs real
% \item Lid-Driven Cavity
%   Show LIC, maybe plot over x?
% \end{itemize}

@@

\section{Accurate Clouds}
We discussed the capabilities of our scheme regarding standard \textsc{cfd} testing scenario.
In this section, we consider our bubble scenarios.
\sidetitle{Cosine Bubble}
\begin{figure}[htb]
  \centering
  \begin{subfigure}[t]{0.5\textwidth}
    \centering
    \includegraphics[scale=1.0]{thesis_cosine_bubble_contour_t700}
    \caption{\label{fig:cosine-bubble-contour-ours}%
      Our result}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.5\textwidth}
    \centering
    \includegraphics{giraldo_results_cosine_bubble}
    \caption{\label{fig:cosine-bubble-contour-giraldo}%
      Figure taken from~\cite{giraldo2008study}, resolution of \SI{20}{\m}, $N = 10$, \dg{}, modified Navier-Stokes} %poly order
  \end{subfigure}
  \caption{\label{fig:cosine-bubble-contour}%
    Contours for the cosine bubble results.
    Contours are $-0.05, -0.025, \ldots 0.525$, excluding zero.
  }
\end{figure}
We ran our first scenario, the cosine bubble, with an \aderdg{}-scheme of order 4 and the adaptive mesh criterion \cref{eq:refinement-criterion} with thresholds of $T_\text{refine} = 2.5$ and $T_\text{delete} = -0.5$.
Our coarsest level has $3^3$ cells, our finest level has $3^4$ cells.
We thus have a resolution of ca.\ \SI{37.037}{\m} up to \SI{12.3457}{\m}.

@@

\begin{figure}[p]
  \vspace*{-4.5cm}
  \thisfloatpagestyle{empty}
  \centering
  \begin{subfigure}{1\textwidth}
    \centering
    \includegraphics{thesis_two_bubbles_contour_collage.pdf}
  \end{subfigure}
  \vspace*{1cm}
  \begin{subfigure}[t]{1\textwidth}
    \centering
    \includegraphics[trim=0.0cm 0.0cm 0.0cm 0.0cm, clip, scale=1.2]{giraldo_results_two_bubbles}
  \end{subfigure}
  \caption{\label{fig:two-bubbles-contour}%
    Results for two bubble scenario.
    On the top is our, the bottom plots are taken from~\cite{muller2010adaptive}.
  Contour values for potential temperature pertubation are $-0.05, 0.05, 0.1, \ldots 0.45$.
}
\end{figure}

@@

\begin{figure}[htb]
  \centering
  \includegraphics{thesis_two_bubbles_contour_t600_noamr} 
  \caption{\label{fig:two-bubbles-noamr-contour}%
    Two bubbles run with fully refined mesh}
\end{figure}
For both settings, we used an \aderdg{}-method of order $4$ and thus have a minimal effective resolution of ca.\ $\SI{3.09}{\m}$.
The final result for the run with the fully refined mesh (\cref{fig:two-bubbles-noamr-contour}) looks identical to the \amr{} results (\cref{fig:two-bubbles-contour}) shown in the previous section.

\sidetitle{Time to solution test}
We now want to compare the time to solution for both settings.
To ensure a fair comparison, we ran both configurations with the same optimisation options:
\begin{verbatim}
"architecture": "hsw",
"shared_memory": {
    "cores": 28,
    "background_job_consumers": 6,
    "properties_file": "sharedmemory.properties",
    "autotuning_strategy": "dummy"
},
"optimisation": {
    "fuse_algorithmic_steps": true,
    "fuse_algorithmic_steps_factor": 0.99,
    "spawn_predictor_as_background_thread": true,
    "spawn_amr_background_threads": true,
    "spawn_update_as_background_thread": true
}
\end{verbatim}
We ran the simulation without plotters to focus on the performance of the numerics and the mesh refinement routines.

@@

We solely used the \tbb{} parallelisation of \exahype{}.
To find a reasonable parameter for the number of background job consumers, we ran a grid search.
We ran three \amr{} testcases for a simulation time of $\SI{1}{\min}$.
The testcases were identical in setting to the aforementioned \amr{} run, but with two levels of refinement.
\todo{Is a plot necessary here?}
The best median performance was achieved by using six background threads.
Note that the results were noisy and it is unclear whether the chosen number of background threads is optimal.

@@

To summarise, our \amr{} criterion saves a significant amount of computational effort while preserving the fine details of a fully refined grid.
\section{Reactive Euler \textit{\&} Navier Stokes}
As a final benchmark, we used the reactive Navier-Stokes equations.
We used a limiting \aderdg{}-scheme with an order of 3.
For runs without viscosity we used a meshsize of $81 \times 81$, for the runs with a viscosity of $\mu = 0.01$ we used a coarser mesh consisting in $27 \times 27$ cells.
We chose the smaller grid because our boundary conditions become unstable for finer grids.
A reason for this might be the fact that we are unable to prescribe the gradient at the boundary in the \muscl{} implementation of \exahype{}.
For some runs, we used the \textsc{dmp} settings
\begin{verbatim}
"limiter": {
    "dmp_observables": 3,
    "dmp_relaxation_parameter": 0.05,
    "dmp_difference_scaling": 1e-3,
    "helper_layers": 1
}
\end{verbatim}
where the \texttt{dmp\_relaxation\_parameter} refers to $\varepsilon_0$ and \texttt{dmp\_difference\_scaling} refers to $\varepsilon$ in \cref{eq:limiting-dmp}.
The large value of $\varepsilon_0$ is necessary, as otherwise the full domain is refined.
We use $\Qrho, \pressure$ and $\QZZ$ as \textsc{dmp} variables.
\begin{figure}[tb]
  \centering
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_euler_nodmp_pressure}
    \caption{Pressure, Euler, no \textsc{dmp}}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_euler_dmp_pressure}
    \caption{Pressure, Euler, \textsc{dmp}}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_ns_pressure}
    \caption{Pressure, \textsc{NS}, no \textsc{dmp}}
  \end{subfigure}

  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_euler_nodmp_Z}
    \caption{$Z$, Euler, no \textsc{dmp}}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_euler_dmp_Z}
    \caption{$Z$, Euler, \textsc{dmp}}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_ns_Z}
    \caption{$Z$, \textsc{NS}, no \textsc{dmp}}
  \end{subfigure}

  \caption{\label{fig:detonation-line-plots}%
    Slices along the $x$-axis for our three configurations.
    We only show one half of the slice because the scenario is symmetric.
  }
\end{figure}
The results for runs with and without \textsc{dmp} can be seen in \cref{fig:detonation-line-plots}.
Note, that oscillations appeared for the Euler equations unless we use the \textsc{dmp}-criterion.
This is not the case for the Navier Stokes equations.
This scenario shows the differences between the viscous Navier-Stokes equations and the Euler equations clearly:
The viscous effects smooth out the otherwise sharp curves.

Furthermore, they have the effect that fewer cells are limited~(\cref{fig:detonation-pressure-contour}).\todo{This is limiter status $\geq$ 3!}
Note that the limiter activation are not symmetric which is unexpected for a radial-symmetric scenario.
The contours for pressure and mass fraction look similar for all considered settings.

It is difficult to compare these results with our references, because they both use different settings.
In constrast to~\cite{helzel2000modified} we use a non-stiff source term and viscous effects,~\cite{hidalgo2011ader} includes viscosity but is limited to one-dimensional scenarios.
We can see that our simulation captures the overall behaviour of the reaction similar to~\cite{helzel2000modified} while the viscous effects have a similar smoothing effect as in~\cite{hidalgo2011ader}.

@@

>>>>

\chapter{Results}\label{chap:results}
We discussed various scenarios in the previous chapter.
In this chapter, we use them to gauge the quality of our numerical scheme.
We start with a convergence test, followed by three standard testing scenarios.
Next, we simulate two different cloud scenarios and check whether our \amr{} criterion is useful.

@@

\section{Accurate CFD}\label{sec:cfd}
After we established that our scheme converges, we evaluate it using standard \textsc{cfd} testing scenarios that either have an analytical solution or for which we have a reference solution.
% Show:
% \begin{itemize}
% \item Taylor-Green High Order
%   LIC, Plot v/p ana vs.\ real
% \item ABC Medium Order
%   Plot p ana.\ vs real
% \item Lid-Driven Cavity
%   Show LIC, maybe plot over x?
% \end{itemize}

@@

\section{Accurate Clouds}\label{sec:results-cloud}
We discussed the capabilities of our scheme regarding standard \textsc{cfd} testing scenario.
In this section, we consider our bubble scenarios.
\sidetitle{Cosine Bubble}
\begin{figure}[htb]
  \centering
  \begin{subfigure}[t]{0.5\textwidth}
    \centering
    \includegraphics[scale=1.0]{thesis_cosine_bubble_contour_t700}
    \caption{\label{fig:cosine-bubble-contour-ours}%
      Our result}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.5\textwidth}
    \centering
    \includegraphics{giraldo_results_cosine_bubble}
    \caption{\label{fig:cosine-bubble-contour-giraldo}%
      Figure taken from~\cite{giraldo2008study}, resolution of \SI{20}{\m}, $N = 10$, \dg{}, modified Navier-Stokes} %poly order
  \end{subfigure}
  \caption{\label{fig:cosine-bubble-contour}%
    Contours for the cosine bubble results.
    Contours are $-0.05, -0.025, \ldots 0.525$, excluding zero.
  }
\end{figure}
We ran our first scenario, the cosine bubble, with an \aderdg{}-scheme of order 4 and the adaptive mesh criterion \cref{eq:refinement-criterion} with thresholds of $T_\text{refine} = 2.5$ and $T_\text{delete} = -0.5$.
Our coarsest level has $3^3$ cells, our finest level has $3^4$ cells.
We thus have a resolution of ca.\ \SI{37.037}{\m} up to \SI{12.3457}{\m}.

@@

% \begin{figure}[p]
%   \vspace*{-4.5cm}
%   \thisfloatpagestyle{empty}
%   \centering
%   \begin{subfigure}{1\textwidth}
%     \centering
%     \includegraphics{thesis_two_bubbles_contour_collage.pdf}
%   \end{subfigure}
%   \vspace*{1cm}
%   \begin{subfigure}[t]{1\textwidth}
%     \centering
%     \includegraphics[trim=0.0cm 0.0cm 0.0cm 0.0cm, clip, scale=1.2]{giraldo_results_two_bubbles}
%   \end{subfigure}
%   \caption{\label{fig:two-bubbles-contour}%
%     Results for two bubble scenario.
%     On the top is our, the bottom plots are taken from~\cite{muller2010adaptive}.
%   Contour values for potential temperature pertubation are $-0.05, 0.05, 0.1, \ldots 0.45$.
% }
% \end{figure}
\begin{figure}[p]
  \thisfloatpagestyle{empty}
  \centering
  \begin{subfigure}[t]{0.5\textwidth}
    \centering
  \includegraphics[trim={0cm 0cm 6cm 0cm},clip]{thesis_two_bubbles_contour_collage_comparison} 
  \caption{Our \amr{} solution.}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.5\textwidth}
    \centering
  % Without z axis on right \includegraphics[trim={8.16cm 0cm 0cm 0cm},clip]{thesis_two_bubbles_contour_collage_comparison} 
  \includegraphics[trim={7.4cm 0cm 0cm 0cm},clip]{thesis_two_bubbles_contour_collage_comparison} 
  \caption{Reference solution from~\cite{muller2010adaptive}.}
  \end{subfigure}
  \caption{\label{fig:two-bubbles-contour}%
     Contours of the potential temperature for two bubble scenario.
     On the left is our solution, the right plots are taken from~\cite{muller2010adaptive}.
     We modified the arrangement of the reference plots and the placement of its axes. 
   Contour values for potential temperature pertubation are $-0.05, 0.05, 0.1, \ldots 0.45$.
  }
\end{figure}

@@

\begin{figure}[htb]
  \centering
  \includegraphics{thesis_two_bubbles_contour_t600_noamr} 
  \caption{\label{fig:two-bubbles-noamr-contour}%
    Two bubbles run with fully refined mesh}
\end{figure}
For both settings, we used an \aderdg{}-method of order $4$ and thus have a minimal effective resolution of ca.\ $\SI{3.09}{\m}$.
The final result for the run with the fully refined mesh (\cref{fig:two-bubbles-noamr-contour}) looks identical to the \amr{} results (\cref{fig:two-bubbles-contour}) shown in the previous section.
Note that the \amr{} criterion manages to track the bubble correctly~(\cref{fig:two-bubbles-amr-grid}) and is introduces only small errors~(\cref{fig:two-bubbles-amr-error})\todo{Increase font size a bit more}.

\begin{figure}[htb]
\centering
\begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{thesis_two_bubbles_amr_grid} 
    \caption{\label{fig:two-bubbles-amr-grid}%
      Potential temperature of \amr{} solution.}
\end{subfigure}\qquad%
\begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{thesis_two_bubbles_amr_error} 
    \caption{\label{fig:two-bubbles-amr-error}%
      Potential temperature of fully refined solution minus \amr{} solution.}
\end{subfigure}

  \caption{%
    \amr{}-grid at $t = \SI{700}{\s}$ for the two bubbles scenario.}
\end{figure}

\sidetitle{Time to solution test}
We now want to compare the time to solution for both settings.
To ensure a fair comparison, we ran both configurations with the same optimisation options:
\begin{verbatim}
"architecture": "hsw",
"shared_memory":

    "cores": 28,
    "background_job_consumers": 6,
    "properties_file": "sharedmemory.properties",
    "autotuning_strategy": "dummy"
},
"optimisation": {
    "fuse_algorithmic_steps": true,
    "fuse_algorithmic_steps_factor": 0.99,
    "spawn_predictor_as_background_thread": true,
    "spawn_amr_background_threads": true,
    "spawn_update_as_background_thread": true
}
\end{verbatim}
We ran the simulation without plotters to focus on the performance of the numerics and the mesh refinement routines.

@@

We solely used the \tbb{} parallelisation of \exahype{}.
To find a reasonable parameter for the number of background job consumers, we ran a grid search.
We ran three \amr{} testcases for a simulation time of $\SI{1}{\min}$.
The testcases were identical in setting to the aforementioned \amr{} run, but with two levels of refinement.
The best median performance was achieved by using six background threads.
Note that the results were noisy and it is unclear whether the chosen number of background threads is optimal.

@@

To summarise, our \amr{} criterion saved a significant amount of computational effort while preserving the fine details of a fully refined grid.
\section{Reactive Euler \textit{\&} Navier Stokes}\label{sec:results-reactive}
As a final benchmark, we used the reactive Navier-Stokes equations.
We used a limiting \aderdg{}-scheme with an order of 3.
For runs without viscosity we used a meshsize of $81 \times 81$, for the runs with a viscosity of $\mu = 0.01$ we used a coarser mesh consisting in $27 \times 27$ cells.
We chose the smaller grid because our boundary conditions become unstable for finer grids.
A reason for this might be the fact that we are unable to prescribe the gradient at the boundary in the \muscl{} implementation of \exahype{}.
For some runs, we used the \textsc{dmp} settings
\begin{verbatim}
"limiter": {
    "dmp_observables": 3,
    "dmp_relaxation_parameter": 0.05,
    "dmp_difference_scaling": 1e-3,
    "helper_layers": 1
}
\end{verbatim}
where the \texttt{dmp\_relaxation\_parameter} refers to $\varepsilon_0$ and \texttt{dmp\_difference\_scaling} refers to $\varepsilon$ in \cref{eq:limiting-dmp}.
The large value of $\varepsilon_0$ is necessary, as otherwise the full domain is refined.
We use $\Qrho, \pressure$ and $\QZZ$ as \textsc{dmp} variables.
\begin{figure}[tb]
  \centering
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_euler_nodmp_pressure}
    \caption{Pressure, Euler, no \textsc{dmp}}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_euler_dmp_pressure}
    \caption{Pressure, Euler, \textsc{dmp}}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_ns_pressure}
    \caption{Pressure, \textsc{NS}, no \textsc{dmp}}
  \end{subfigure}

  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_euler_nodmp_Z}
    \caption{$Z$, Euler, no \textsc{dmp}}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_euler_dmp_Z}
    \caption{$Z$, Euler, \textsc{dmp}}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics{thesis_detonation_ns_Z}
    \caption{$Z$, \textsc{NS}, no \textsc{dmp}}
  \end{subfigure}
  \caption{\label{fig:detonation-line-plots}%
    Slices along the $x$-axis for our three configurations.
    We only show one half of the slice because the scenario is symmetric.
  }
\end{figure}
The results for runs with and without \textsc{dmp} can be seen in \cref{fig:detonation-line-plots}.
Note, that oscillations appeared for the Euler equations unless we use the \textsc{dmp}-criterion.
This is not the case for the Navier Stokes equations.
This scenario shows the differences between the viscous Navier-Stokes equations and the Euler equations clearly:
The viscous effects smooth out the otherwise sharp curves.

Furthermore, they have the effect that fewer cells are limited~(\cref{fig:detonation-pressure-contour}).
Note that the limiter activation are not symmetric which is unexpected for a radial-symmetric scenario.
The contours for pressure and mass fraction look similar for all considered settings.

It is difficult to compare these results with our references, because they both use different settings.
In constrast to~\cite{helzel2000modified} we use a non-stiff source term and viscous effects,~\cite{hidalgo2011ader} includes viscosity but is limited to one-dimensional scenarios.
Comparing with \cref{fig:detonation-contour-reference} we can see that our simulation captures the overall behaviour of the reaction similar to the stiff simulations of~\cite{helzel2000modified}.
The viscous effects have a similar smoothing effect as in~\cite{hidalgo2011ader}.

@@

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.33\textwidth]{leveque_detonation_mass_fraction}
    \caption{\label{fig:detonation-contour-reference}Reference solution for mass fraction with stiff source.
      Image taken from~\cite{helzel2000modified}.}
\end{figure}
